---
title: "<h1 style='text-align:center;'> Common epidemiology measures </h1>"
author: "Hugo Marthinet"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
  toc: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Common epidemiology measures}
  %\usepackage[utf8]{inputenc}
---

\setmainfont{Calibri Light}

```{r, echo = FALSE, message = FALSE}

# If you want to create a PDF document paste the following after line 9 above:
#   pdf_document:
#     toc: true
#     highlight: tango
#     number_sections: no
#     latex_engine: xelatex    
# header-includes: 
#    - \usepackage{fontspec}

knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(epiplus)
```

This package provides calculation functions for commons epidemiology and pharmaco-epidemiology measures. This is merely an easy and accessible way to apply well known formulas.


## Rates and confusion matrix calculus

First use of this package is to compute rates along with the confidence intervals. This can be incidence or mortality rates, standardized or not.

```{r, message = FALSE}
# Parameters
pop_tot <- c(32648555, 32587453) # Total population size
nb_ref <- c(61904, 59910) # Number of cases

MR_ref <- c(190, 183) # Reference mortality rates
pop_obs <- c(300340, 301352) # Studied population size
nb_exp <- c(570, 551) # Expected number of deaths
nb_obs <- c(200, 213) # Observed number of deaths

# Incidence rates
IR(nb_ref, pop_tot)
# Mortality rates (exactly the same, parameters names only change for convenience)
MR(nb_ref, pop_tot)

# Standardized Mortality Rate
SMR(nb_obs, nb_exp, decimal = 3)
SMR(n.obs = nb_obs, n.ref = nb_ref, pop.obs = pop_obs, pop.ref = pop_tot, decimal = 3)
SMR(n.obs = nb_obs, pop.obs = pop_obs, MR.ref = MR_ref, decimal = 3)
```

One can also compute at once every metrics associated to a two by two confusion matrix :

```{r, message = FALSE}
(tab <- matrix(c(67, 2, 4, 7),
                nrow = 2, byrow = TRUE))
confusionMetrics(tab = tab)
```

## Reliability

Inter-raters reliability measures such as Cohen's or Fleiss' Kappas are used to asses the degree of agreement among independent observers who rate, code, or assess the same phenomenon.
Those measures are is bounded between -1 and 1. A negative value corresponds to a poor agreement between raters, whereas a value of 1 correspond to a complete agreement.

```{r, message = FALSE}
(tab <- matrix(c(15, 17, 8, 6),
                nrow = 2, byrow = TRUE))
kappa_cohen(tab)
kappa_fleiss(tab)
```


## (Co)morbidity

Morbidity (Mortality-Related Morbidity Index and Expenditure-Related Morbidity Index) and comorbidity (Charlson Comorbidity Index) indexes are based on comorbidities
observes in the patients. Those consist in a the sum of scores associated to the presence/absence of a given disease or condition. An higher score is associated
to an higher risk.

```{r, message = FALSE}
morbidity(gender = "M", age = 75,
          isch.heart = 0, cerebro.vasc = 0, heart.fail = 0, periph.vasc = 0, diabetes = 1,
          cancer = 0, cancer.hist = 0, schizo = 0, depression = 0, subst.abuse = 0,
          dementia = 0, parkinson = 0,  mult.sclero = 0, epilespy = 0,
          chronic.resp = 1, infl.bowel = 0, rheum.arthr = 0, HIV = 0, es.renal = 0, liver.pancr = 0)

comorbidity(age = 75,
  myocard.infarct = 0, heart.fail = 0, periph.vasc = 0, cerebr = TRUE, diabetes = 0,
  dementia = 0, chronic.pulm = 0, connect.tissue = 0, ulcer = 0, mild.liver = 0,
  hemiplegia = 0, renal = TRUE, diabetes metast.tumor = 0, aids = 0)
```


## Adherence to care

Adherence measures track how well patients are following their treatment. There are many ways to measures adherence, not taking into account the same metrics and making more 
or less assumptions about how the patients are actually using their medications. One must study the literature first, and have good knowladge of the data available, before choosing a specific measure.

```{r, message = FALSE}
# Example for a patient with a care sequence like : ABABBCBD
# A, B and C being diffrent practionners
# Order from most to leat recent (A is the most recent visit)

seq <- "ABABBCBD"
pract <- c("A", "B", "C", "D")
vi <- rep(NA, 4)
for (i in 1:4) {
  vi[i] <- length(lapply(gregexpr(pract[i], seq), attributes)[[1]]$match.length)
}
vi

CoC(w = 8,
    n = 4, # A, B, C and D
    vi = vi,
    vc = 2, # 2 times A
    vu = 4, # 4 times B
    vs = 4) # Assuming every sites are different
```
